---
title: "vAMPirus ASV Report `r commandArgs(trailingOnly=T)[1]`"
date: "Generated on: `r Sys.time()`"
output: html_document
params:
  interactive: TRUE
  reads: !r commandArgs(trailingOnly=T)[2]    # reads
  counts: !r commandArgs(trailingOnly=T)[3]   # csv
  metadata: !r commandArgs(trailingOnly=T)[4] # metadata
  filter: !r commandArgs(trailingOnly=T)[5]   # filter min counts
  heatmap: !r commandArgs(trailingOnly=T)[6]  # heatmap
  tax: !r commandArgs(trailingOnly=T)[7]      # tax
  try: !r commandArgs(trailingOnly=T)[8]      # trymax
---
<style>
.rectangle {
  height: 37px;
  width: 105%;
  background-color: #576675;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
      message = FALSE,
      warning = FALSE,
      out.width="100%")
```
[](https://sync.palmuc.org/index.php/s/gscg4m4PtdN5ZwY/preview)

```{r load_libraries, include=FALSE}
library(vegan)
#library(rstatix)
library(tidyverse)
library(scales)
library(cowplot)
library(dplyr)
library(ggtree)
library(plotly)
library(BiocParallel)
library(knitr)
library(kableExtra) #install.packages("kableExtra")
library(rmarkdown)
library(processx) #install.packages("processx")
register(MulticoreParam(4))
```

```{r colors, include=FALSE}
mycol=c('#088da5','#73cdc8','#ff6f61','#7cb8df','#88b04b','#00a199','#6B5B95','#92A8D1','#b0e0e6','#ff7f50','#088d9b','#E15D44','#e19336')
```
<br>

NOTE: Most plots are interactive and you can use the legend to specify samples/treatment of interest

<br>
<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Pre- and Post-Adapter Removal Read Stats</h2></div>
<br>
```{r readstats, echo=FALSE}
reads_stats=read.csv(params$reads)
#reads_stats=read.csv("PVID_final_reads_stats.csv")
paged_table(reads_stats,options = list(rows.print = 20))

#test plotly table
fig <- plot_ly(
    type = 'table',
    header = list(values=names(reads_stats),
                  align = c('left'),
                  line = list(width = 1, color = 'black'),
                  fill = list(color = 'rgb(45, 100, 230)')),
    cells = list(values=unname(reads_stats),
                 line = list(width = 1, color = 'black'),
                 fill = list(color = 'white'),
                 align = c('center'),
                 font = list(color = '#506784', size = 10))
    )

#fig
```
<br>
<br>


### Total number of reads before and after adapter removal

```{r readstats_plot, echo=FALSE}
# Plot of reads before and after
ptotal <- plot_ly(typle="box",marker=list(colors=mycol))
ptotal <- ptotal %>% add_boxplot(y=reads_stats$Total_before, name="Reads before filtering")
ptotal <- ptotal %>% add_boxplot(y=reads_stats$Total_after, name="Reads after filtering")
#ptotal <- ptotal %>% layout(title=list(text="Number of reads before and after filtering"))
ptotal <- ptotal %>% layout(legend = list(x=10,y=.5))
ptotal
```
<br>

### Forward (R1) and reverse (R2) read length before and after adapter removal

```{r readstats_plot2, echo=FALSE}
# Plot of R1 and R2 before and after
pr <- plot_ly(y=reads_stats$R1_before_length, type="box", name="R1 length before")
pr <- pr %>% add_boxplot(y=reads_stats$R1_after_length, name="R1 length after")
pr <- pr %>% add_boxplot(y=reads_stats$R2_before_length, name="R2 length before")
pr <- pr %>% add_boxplot(y=reads_stats$R2_after_length, name="R2 length after")
#pr <- pr %>% layout(title = "R1 and R2 Length")
pr <- pr %>% layout(legend = list(x=10,y=.5))
pr
```
<br>
<br>
<br>
```{r load_datasets, include=FALSE}
sample_name=params$counts
sample_metadata=params$metadata
#sample_name="PVID_vAMPtest1_otu.85_counts.csv"
#sample_metadata="PVID_fiscesTestmeta.csv"
#sample_name="vAMPrun_otu.55_counts.csv"
#sample_metadata="pvid_samples.csv"
data<- read.csv(sample_name, check.names=FALSE)
data2 <-as.data.frame(t(data))
data2$sample <- row.names(data2)
colnames(data2)<- as.matrix(data2[1,])
as.data.frame(data2)
data2 <- data2[-1,]

#X.OTU.ID for X.Sequence.
data2 <- data2 %>%
  rename(sample=OTU_ID)
data2dim <- dim(data2)

##Loading metadata
samples <- read.csv(sample_metadata, header=TRUE)

##Combining data and metadata
data3 <- merge(data2, samples, by="sample")

dim_data3 <- dim(data3)
dim_samples <- dim(samples)
cols <- dim_data3[2]-dim_samples[2]+1
first <-colnames(data3)[2]
last <- colnames(data3)[cols]
data3[,2:cols] <- lapply(data3[,2:cols], as.character)
data3[,2:cols] <- lapply(data3[,2:cols], as.numeric)

#Calculate total reads per sample
data4 <- data3%>%
  mutate(sum=select(.,2:cols)%>%
           apply(1, sum, na.rm=TRUE))

```


<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Number of Reads Per Sample</h2></div>

```{r plot, echo=FALSE}
# sample and count
con <- plot_ly(data4, x = ~sum, y = ~sample, name = "Sample", type = 'scatter',
               mode = "markers", marker = list(color = "#088da5"), hovertemplate = paste('Sample: %{y}','<br>Total reads: %{x}','<extra></extra>'))
con <- con %>% layout(xaxis = list(title = "Total reads"),yaxis = list(title = "Sample"))
con
```
<br>
<br>
<br>
```{r filter_data, include=FALSE}
##Filter samples with low reads
nfil=params$filter
data5 <- data4 %>%
  filter(sum>nfil)
   #can cause errors
data5dim <-dim(data5)
minreads<-min(data5$sum)
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Rarefaction</h2></div>
```{r rarefaction, echo=FALSE, cache=FALSE}
##Rarefaction curves
rarefaction <- rarecurve(data5[,2:cols])

##rarefied dataset
raredata <- as.data.frame(rrarefy(data5[,2:cols], sample=minreads))
```
<br>
<br>
<br>
<br>
<br>


<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Diversity Analyses Plots</h2></div>

<br>
<br>

### Shannon diversty

<br>
```{r diversity_analysis, echo=FALSE}

index <-diversity(raredata, index= "shannon")
shannondata5 <- as.data.frame(index)

index <- diversity(raredata, index= "simpson")
simpsondata5 <- as.data.frame(index)

mind5<-min(data5$sum)
index <- rarefy(data5[,2:cols], sample=mind5)
rarerichnessdata5 <- as.data.frame(index)

metadata <- data5[,(cols+1):data5dim[2]]

metadata$sample <- data5$sample
shannondata5$sample<- data5$sample
simpsondata5$sample<- data5$sample
rarerichnessdata5$sample <-data5$sample

shannondata5_2 <- merge(shannondata5, metadata, by="sample")
simpsondata5_2 <- merge(simpsondata5, metadata, by="sample")
rarerichnessdata5_2 <- merge(rarerichnessdata5, metadata, by="sample")

#shannonplot <- ggplot(shannondata5_2, aes(x=treatment,color=treatment,y=index))+
#  geom_boxplot()+
#  geom_point()+
#  theme_classic()+
#  labs(y="Index",x="Treatment")+
#  theme(axis.text=element_text(size=12))+
#  theme(legend.position = "none")

#sh<-ggplotly(shannonplot)
#sh <- sh %>% layout(title = list(text="Shannon diversty",y=.99))
#sh

sh <- plot_ly(shannondata5_2, x=~treatment, y=~index, color=~treatment, colors=mycol, type="box", boxpoints = "all", pointpos = 0, jitter = 0.5)
#sh <- sh %>% layout(title = list(text="Shannon diversty",y=.99))
sh <- sh %>% layout(legend = list(x=10,y=.5))
sh <- sh %>% layout(yaxis=list(title = "Index"))
sh <- sh %>% layout(xaxis=list(title = "Treatment"))
sh
```
<br>
<br>
<br>
<br>

### Simpson diversty

<br>
```{r diversity_analysis2, echo=FALSE}
#simpsonplot <- ggplot(simpsondata5_2, aes(x=treatment,color=treatment,y=index))+
#  geom_boxplot()+
#  geom_point()+
#  theme_classic()+
#  labs(y="Index",x="Treatment")+
#  theme(axis.text=element_text(size=12))+
#  theme(legend.position = "none")

#s<-ggplotly(simpsonplot)
#s <- s %>% layout(title = list(text="Simpson diversty",y=.99))
#s

s <- plot_ly(simpsondata5_2, x=~treatment, y=~index, color=~treatment, colors=mycol, type="box", boxpoints = "all", pointpos = 0, jitter = 0.5)
#s <- s %>% layout(title = list(text="Simpson diversty",y=.99))
s <- s %>% layout(legend = list(x=10,y=.5))
s <- s %>% layout(yaxis=list(title = "Index"))
s <- s %>% layout(xaxis=list(title = "Treatment"))
s
```
<br>
<br>
<br>
<br>

### Species Richness

<br>
```{r diversity_analysis3, echo=FALSE}
#richnessplot <- ggplot(rarerichnessdata5_2, aes(x=treatment,color=treatment,y=index))+
#  geom_boxplot()+
#  geom_point()+
#  theme_classic()+
#  labs(y="Richness",x="Treatment")+
#  theme(axis.text=element_text(size=12))+
#  theme(legend.position = "none")

ri <- plot_ly(rarerichnessdata5_2, x=~treatment, y=~index, color=~treatment, colors=mycol, type="box", boxpoints = "all", pointpos = 0, jitter = 0.5)
#ri <- ri %>% layout(title = list(text="ASV Richness",y=.99))
ri <- ri %>% layout(legend = list(x=10,y=.5))
ri <- ri %>% layout(yaxis=list(title = "Index"))
ri <- ri %>% layout(xaxis=list(title = "Treatment"))
ri
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Distance To Centroid</h2></div>
<br>
```{r distance, echo=FALSE}
##Distance
intermediate <- raredata
bray <- vegdist(sqrt(intermediate), method="bray")

##Dispersion
disper <- betadisper(bray, group = metadata$treatment, type="centroid")
df <- data.frame(Distance_to_centroid=disper$distances,Group=disper$group)
df$sample <- data5$sample
df2 <- merge(df, metadata, by="sample")

#p<- ggplot(data=df2,aes(x=treatment,y=Distance_to_centroid,colour=treatment))+
#  geom_boxplot(outlier.alpha = 0)+
#  theme_classic()+
#  geom_point(position=position_dodge(width=0.75))+
#  labs(y="Distance",x="Treatment")+
#  theme(axis.text=element_text(size=12))+
#  theme(legend.position = "none")

#cen <- ggplotly(p)
#cen <- cen %>% layout(title = list(text="Distance to centroid",y=.99))
#cen

cen <- plot_ly(df2, x=~treatment, y=~Distance_to_centroid, color=~treatment, colors=mycol, type="box", boxpoints = "all", pointpos = 0, jitter = 0.5)
#cen <- cen %>% layout(title = list(text="Distance to centroid",y=.99))
cen <- cen %>% layout(legend = list(x=10,y=.5))
cen <- cen %>% layout(yaxis=list(title = "Distance"))
cen <- cen %>% layout(xaxis=list(title = "Treatment"))
cen
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;NMDS Plots</h2></div>

<br>

### 2D NMDS

<br>
```{r nmds2d, echo=FALSE}
##NMDS
datax <- decostand(raredata,method="total") #method 'total' normalizes data to sum up to 1 --data5[,2:cols]

MDS <- metaMDS(sqrt(datax),
               distance = "bray",autotransform = FALSE,
               k = 2,
               maxit = 999,
               trymax = params$try,
               wascores = TRUE)

if (MDS$converged == "TRUE") {

data.scores <- as.data.frame(scores(MDS))
data.scores$sample <- data5$sample
data.scores.2 <- merge(data.scores, metadata, by="sample")

p <- ggplot(data.scores.2, aes(x=NMDS1, y=NMDS2,color=treatment))+
  geom_point(size=2)+
  theme_classic()+
  theme(legend.title = element_blank())

#fff <- ggplotly(p)
#fff <- fff %>% layout(legend=list(y=.5))
#fff

fff <-plot_ly(data.scores.2, x=~NMDS1, y=~NMDS2, color=~treatment, colors=mycol, text = ~paste("Sample: ", sample))
fff <- fff %>% layout(legend=list(y=.5))
fff

} else {
  print("No Convergence")
}
```

<br>
<br>
<br>
<br>

### 3D NMDS

<br>

```{r nmds3d, echo=FALSE}
MDS3 <- metaMDS(sqrt(datax),
               distance = "bray",autotransform = FALSE,
               k = 3,
               maxit = 999,
               trymax = params$try,
               wascores = TRUE)

if (MDS3$converged == "TRUE") {

data.scores3 <- as.data.frame(scores(MDS3))
data.scores3$sample <- data5$sample
data.scores.3 <- merge(data.scores3, metadata, by="sample")
p3d <- plot_ly(data.scores.3, x= ~ NMDS1, y = ~ NMDS2, z = ~ NMDS3, text = ~paste("Sample: ", sample),
        color = ~treatment, colors = mycol,
        mode = 'markers', symbol = ~treatment, symbols = c('square','circle'),
        marker = list(opacity = .8,line=list(color = 'darkblue',width = 1))
        )
p3d <- p3d %>% layout(legend=list(y=.5))
p3d

} else {
  print("No Convergence")
}
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;OTU Abundance Per Sample</h2></div>

```{r long, echo=FALSE}
dataz <- decostand(data5[,2:cols],method="total") #method 'total' normalizes data to sum up to 1
dataz$sample <- data5$sample
datay <- merge(dataz, metadata, by="sample")
datalong <- datay %>%
  tidyr::gather(first:last, key=hit, value=reads)

##Barplot
## add better colors
#spec_bar <- ggplot(datalong, aes(x=forcats::fct_reorder(timepoint, as.numeric(as.character(timepoint))),
#    y=reads, fill=hit))+ #Consider sqrt-transforming data
#    geom_bar(aes(), stat="identity", position="fill")+
#    #coord_polar("y", start=0)+
#    theme_classic()+
#    facet_wrap(colony~treatment, nrow=2)+
#    #coord_flip()+
#    labs(x="Timepoint") #+ theme(legend.position = "none")
#spec_bar <- ggplot(datalong, aes(x=sample,y=reads,fill=hit))+geom_bar(aes(), stat="identity", position="fill")+
#  theme(axis.text.x=element_text(angle=90))
#ggplotly(spec_bar)
#,'#00b300','#00b3b3','#0059b3','#6600ff','#b800e6','#ff3333','#ff8000','#ffff00','#bf8040','#42bcf5','#b428f5','#2e8d7e','#664e7e','#a4c700','#1aa3ff'

ddd <- plot_ly(datalong, x=~sample, y=~reads, color=~hit, colors=mycol)
ddd <- ddd %>% layout(type='bar',barmode = 'stack')
ddd <- ddd %>% layout(yaxis=list(title = "Relative abundance"))
ddd <- ddd %>% layout(xaxis=list(title = "Sample"))
ddd <- ddd %>% layout(legend = list(x=10,y=.5))
ddd
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;OTU Abundance Per Treatment</h2></div>

```{r asv_barplot, echo=FALSE}
datalong <- datalong %>%
  filter(reads>0)

#asv_bar <- ggplot(datalong, aes(x=reorder(hit,reads), y=reads, fill=treatment))+
#  geom_bar(stat="identity")+
#  scale_fill_manual(values=c('#088da5','#e19336'))+ #colors not working
#  coord_flip()+
#  theme_classic()+
#  theme(axis.title.y = element_blank())+
#  theme(legend.title = element_blank())

asp2 <- plot_ly(datalong, y=~hit, x=~reads, color=~treatment, colors=mycol,text = ~paste("Sample: ", sample), opacity=.9)
asp2 <- asp2 %>% layout(type='bar', barmode = 'group')
asp2 <- asp2 %>% layout(yaxis = list(title = '', categoryorder = "total ascending"))
asp2 <- asp2 %>% layout(legend = list(x=10,y=.5))
asp2
```
<br>
<br>
<br>
<br>
<br>

<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Pairwise Percent-ID Heatmap</h2></div>

<br>
<br>
```{r heatmap, echo=FALSE}
heatdata=params$heatmap
#heatdata="a.csv"
#heatdata="PVID_vAMPtest1_otu.85_PercentID.matrix"
simmatrix<- read.csv(heatdata, header=FALSE)
rownames(simmatrix) <- simmatrix[,1]
simmatrix <- simmatrix[,-1]
colnames(simmatrix) <-rownames(simmatrix)
cols <- dim(simmatrix)[2]
simmatrix$AA <- rownames(simmatrix)
rval=nrow(simmatrix)
simmatrix2 <- simmatrix %>%
  gather(1:rval, key=sequence, value=similarity)
x=reorder(simmatrix2$AA,simmatrix2$similarity)
y=reorder(simmatrix2$sequence,simmatrix2$similarity)
similaritymatrix <- ggplot(simmatrix2, aes(x=x, y=y,fill=similarity))+
  geom_raster()+
  scale_fill_distiller(palette="Spectral")+
  theme(axis.text.x = element_text(angle = 90))+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())


ggplotly(similaritymatrix)
```
<br>
<br>
<br>
<br>
<br>


<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Taxonomy Result Visualization</h2></div>

<br>
<br>

```{r taxonomy, echo=FALSE}
tax=read.csv(params$tax,header=F)
#tax=read.csv("tax.csv",header=F)
#tax=read.csv("PVID_vAMPtest1_otu.85_summary_for_plot.csv", header=F)
tp <- plot_ly(tax, labels = ~V1, values = ~V2)
tp <- tp %>% add_pie(marker=list(colors=mycol, line=list(color='#000000', width=.5)), hole = 0.6)
tp <- tp %>% layout(title = "Taxonomy distribution",  showlegend = F,
                    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
                    )

tp
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
