---
title: "vAMPirus Data Check Report: `r commandArgs(trailingOnly=T)[1]`"
date: "Generated on: `r Sys.time()`"
output: html_document
params:
  interactive: TRUE
  fastpcsv: !r commandArgs(trailingOnly=T)[2]
  reads_per_sample_preFilt: !r commandArgs(trailingOnly=T)[3]
  read_per_sample_postFilt: !r commandArgs(trailingOnly=T)[4]
  preFilt_baseFrequency: !r commandArgs(trailingOnly=T)[5]  
  postFilt_baseFrequency: !r commandArgs(trailingOnly=T)[6]
  preFilt_qualityScore: !r commandArgs(trailingOnly=T)[7]
  postFilt_qualityScore: !r commandArgs(trailingOnly=T)[8]
  preFilt_averageQuality: !r commandArgs(trailingOnly=T)[9]
  postFilt_averageQuaulity: !r commandArgs(trailingOnly=T)[10]
  preFilt_length: !r commandArgs(trailingOnly=T)[11]
  postFilt_length: !r commandArgs(trailingOnly=T)[12]
  number_per_percentage_nucl: !r commandArgs(trailingOnly=T)[13]
  number_per_percentage_prot: !r commandArgs(trailingOnly=T)[14]
---
<style>
.rectangle {
  height: 37px;
  width: 105%;
  background-color: #576675;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
      message = FALSE,
      warning = FALSE,
      out.width="100%")
```

```{r load_libraries, include=FALSE}
library(vegan)
#library(rstatix)
library(tidyverse)
library(scales)
library(cowplot)
library(dplyr)
library(ggtree)
library(plotly)
library(BiocParallel)
library(knitr)
library(kableExtra) #install.packages("kableExtra")
library(rmarkdown)
register(MulticoreParam(4))
```

```{r colors, include=FALSE}
mycol=c('#088da5','#73cdc8','#ff6f61','#7cb8df','#88b04b','#00a199','#6B5B95','#92A8D1','#b0e0e6','#ff7f50','#088d9b','#E15D44','#e19336')
```
<br>
<br>
<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Pre- and Post-Adapter Removal Read Stats</h2></div>
<br>
```{r readstats, echo=FALSE}
reads_stats=read.csv(params$fastpcsv)
#reads_stats=read.csv("final_reads_stats.csv")
paged_table(reads_stats,options = list(rows.print = 20))
```
<br>
<br>

### Total number of reads before and after adapter removal

```{r readstats_plot, echo=FALSE}
# Plot of reads before and after
ptotal <- plot_ly(typle="box",marker=list(colors=mycol))
ptotal <- ptotal %>% add_boxplot(y=reads_stats$Total_before, name="Reads before filtering")
ptotal <- ptotal %>% add_boxplot(y=reads_stats$Total_after, name="Reads after filtering")
#ptotal <- ptotal %>% layout(title=list(text="Number of reads before and after filtering"))
ptotal <- ptotal %>% layout(legend = list(x=10,y=.5))
ptotal
```
<br>

### Forward (R1) and reverse (R2) read length before and after adapter removal

```{r readstats_plot2, echo=FALSE}
# Plot of R1 and R2 before and after
pr <- plot_ly(y=reads_stats$R1_before_length, type="box", name="R1 length before")
pr <- pr %>% add_boxplot(y=reads_stats$R1_after_length, name="R1 length after")
pr <- pr %>% add_boxplot(y=reads_stats$R2_before_length, name="R2 length before")
pr <- pr %>% add_boxplot(y=reads_stats$R2_after_length, name="R2 length after")
#pr <- pr %>% layout(title = "R1 and R2 Length")
pr <- pr %>% layout(legend = list(x=10,y=.5))
pr
```
<br>
<br>
<br>
<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Post-Merging Read Stats</h2></div>
<br>
<br>

### Pre-filtering reads per sample

<br>

```{r preFilt_preClean_reads, echo=FALSE}
prer=read.csv(params$reads_per_sample_preFilt, header=TRUE)
prerp <- plot_ly(prer, x = ~reads, y = ~sample, name = "Sample", type = 'scatter',
               mode = "markers", marker = list(color = "#088da5"),
               hovertemplate = paste('Sample: %{y}','<br>Total reads: %{x}','<extra></extra>'))
prerp <- prerp %>% layout(xaxis = list(title = "Total reads"),yaxis = list(title = "Sample"))
prerp
```

<br>

### Post-filtering reads per sample

<br>

```{r postFilt_postClean_reads, echo=FALSE}
postr=read.csv(params$read_per_sample_postFilt, header=TRUE)
postrp <- plot_ly(postr, x = ~reads, y = ~sample, name = "Sample", type = 'scatter',
               mode = "markers", marker = list(color = "#088da5"),
               hovertemplate = paste('Sample: %{y}','<br>Total reads: %{x}','<extra></extra>'))
postrp <- postrp %>% layout(xaxis = list(title = "Total reads"),yaxis = list(title = "Sample"))
postrp
```

<br>

### Pre-filtering base frequency per position on reads

<br>

```{r preClean_baseFrequency, echo=FALSE}
prec=read.csv(params$preFilt_baseFrequency, header=TRUE)
#prec=read.csv("PVID_vAMPtest1_all_merged_preFilt_preClean_baseFrequency_hist.csv", header=TRUE)
precp <- plot_ly(x=c(1:nrow(prec)),y=as.numeric(prec[,2]), type="scatter", mode = "markers", name="A", opacity = .8, markers=list(color='#088da5'))
precp <- precp %>% add_trace(y=as.numeric(prec[,3]), name="C", opacity = .8, markers=list(color='#e19336'))
precp <- precp %>% add_trace(y=as.numeric(prec[,3]), name="G", opacity = .8, markers=list(color='#ff6f61'))
precp <- precp %>% add_trace(y=as.numeric(prec[,3]), name="T", opacity = .8, markers=list(color='#7cb8df'))
precp <- precp %>% add_trace(y=as.numeric(prec[,3]), name="N", opacity = .8, markers=list(color='#88b04b'))
precp <- precp %>% layout(xaxis=list(title="Base position"))
precp <- precp %>% layout(yaxis=list(title="Base Frequency", range = c(0,1)))
precp <- precp %>% layout(legend = list(x=10,y=.5), hovermode = "x unified")
precp
```

<br>

### Post-filtering base frequency per position on reads

<br>

```{r postClean_baseFrequency, echo=FALSE}
postc=read.csv(params$postFilt_baseFrequency, header=TRUE)
#postc=read.csv("PVID_vAMPtest1_all_merged_postFilt_baseFrequency_hist.csv", header=TRUE)
postcp <- plot_ly(x=c(1:nrow(postc)),y=as.numeric(postc[,2]),type="scatter", mode = "markers", name="A", opacity = .8, markers=list(color='#088da5'))
postcp <- postcp %>% add_trace(y=as.numeric(postc[,3]), name="C", opacity = .8, markers=list(color='#e19336'))
postcp <- postcp %>% add_trace(y=as.numeric(postc[,3]), name="G", opacity = .8, markers=list(color='#ff6f61'))
postcp <- postcp %>% add_trace(y=as.numeric(postc[,3]), name="T", opacity = .8, markers=list(color='#7cb8df'))
postcp <- postcp %>% add_trace(y=as.numeric(postc[,3]), name="N", opacity = .8, markers=list(color='#88b04b'))
postcp <- postcp %>% layout(xaxis=list(title="Base position"))
postcp <- postcp %>% layout(yaxis=list(title="Base Frequency", range = c(0,1)))
postcp <- postcp %>% layout(legend = list(x=10,y=.5), hovermode = "x unified")
postcp
```

<br>

### Pre-filtering mean quality score per position on reads

<br>

```{r preClean_qualityScore, echo=FALSE}
preq=read.csv(params$preFilt_qualityScore, header=TRUE)
preqp <- plot_ly(x=c(1:nrow(preq)),y=as.numeric(preq[,2]),type="scatter", mode = "lines", name="Reads", opacity = .8, line=list(color='#088da5'))
preqp <- preqp %>% layout(xaxis=list(title="Base position"))
preqp <- preqp %>% layout(yaxis=list(title="Mean quality", range = c(0,41)))
preqp <- preqp %>% layout(legend = list(x=10,y=.5), hovermode = "x unified")
preqp
```

<br>

### Post-filtering mean quality score per position on reads

<br>

```{r postClean_qualityScore, echo=FALSE}
postq=read.csv(params$postFilt_qualityScore, header=TRUE)
#postq=read.csv("PVID_vAMPtest1_all_merged_postFilt_qualityScore_hist.csv", header=TRUE)
postqp <- plot_ly(x=c(1:nrow(postq)),y=as.numeric(postq[,2]),type="scatter", mode = "lines", name="Reads", opacity = .8, line=list(color='#088da5'))
postqp <- postqp %>% layout(xaxis=list(title="Base position"))
postqp <- postqp %>% layout(yaxis=list(title="Mean quality", range = c(0,41)))
postqp <- postqp %>% layout(legend = list(x=10,y=.5), hovermode = "x unified")
postqp
```

<br>

### Pre-filtering read GC-content

<br>

```{bash gc, echo=FALSE}
cat *all_merged_preFilt_preClean_gcContent_hist.csv | grep -v "#" >final_gc.csv
cat *all_merged_preFilt_preClean_gcContent_hist.csv | grep "#" | tr -d "#"
```
```{r gc_plot, echo=FALSE}
pregc=read.csv("final_gc.csv", header=F)
pregcp <- plot_ly(pregc, x=~V1, y=~V2, type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('GC content: %{x}','<br>Count: %{y}','<extra></extra>'))
pregcp <- pregcp %>% layout(yaxis = list(title = "Count"))
pregcp <- pregcp %>% layout(xaxis = list(title = "GC content"))
pregcp
```

<br>

### Post-filtering read GC-content

<br>

```{bash gc2, echo=FALSE}
cat *all_merged_postFilt_gcContent_hist.csv | grep -v "#" >final_gc2.csv
cat *all_merged_postFilt_gcContent_hist.csv | grep "#" | tr -d "#"
```
```{r gc2_plot, echo=FALSE}
postgc=read.csv("final_gc2.csv", header=F)
postgcp <- plot_ly(postgc, x=~V1, y=~V2, type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('GC content: %{x}','<br>Count: %{y}','<extra></extra>'))
postgcp <- postgcp %>% layout(yaxis = list(title = "Count"))
postgcp <- postgcp %>% layout(xaxis = list(title = "GC content"))
postgcp
```


<br>

### Number of reads per quality score (pre-filtering)

<br>


```{r pre_averageQuality_hist, echo=FALSE}
preaq=read.csv(params$preFilt_averageQuality, header=TRUE)
#preaq=read.csv("PVID_vAMPtest1_all_merged_preFilt_preClean_averageQuality_hist.csv", header=T)
preaqp <- plot_ly(preaq, x=preaq[,1], y=preaq[,2], type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('Average quality: %{x}','<br>Count: %{y}','<extra></extra>'))
preaqp <- preaqp %>% layout(yaxis = list(title = "Count"))
preaqp <- preaqp %>% layout(xaxis = list(title = "Average quality"))
preaqp
```



<br>

### Number of reads per quality score (post-filtering)

<br>


```{r post_averageQuality_hist, echo=FALSE}
postaq=read.csv(params$postFilt_averageQuaulity, header=TRUE)
#postaq=read.csv("PVID_vAMPtest1_all_merged_postFilt_averageQuaulity_hist.csv", header=T)
postaqp <- plot_ly(postaq, x=postaq[,1], y=postaq[,2], type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('Average quality: %{x}','<br>Count: %{y}','<extra></extra>'))
postaqp <- postaqp %>% layout(yaxis = list(title = "Count"))
postaqp <- postaqp %>% layout(xaxis = list(title = "Average quality"))
postaqp
```


<br>

### Number of reads per length (pre-filtering)

<br>


```{r pre_length_hist, echo=FALSE}
prelh=read.csv(params$preFilt_length, header=TRUE)
#prelh=read.csv("PVID_vAMPtest1_all_merged_preFilt_preClean_length_hist.csv", header=T)
prelhp <- plot_ly(prelh, x=prelh[,1], y=prelh[,2], type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('Read Length: %{x}','<br>Count: %{y}','<extra></extra>'))
prelhp <- prelhp %>% layout(yaxis = list(title = "Count"))
prelhp <- prelhp %>% layout(xaxis = list(title = "Read Length"))
prelhp
```

<br>

### Number of reads per length (post-filtering)

<br>

```{r post_length_hist, echo=FALSE}
postlh=read.csv(params$postFilt_length, header=TRUE)
#postlh=read.csv("PVID_vAMPtest1_all_merged_postFilt_length_hist.csv", header=T)
postlhp <- plot_ly(postlh, x=postlh[,1], y=postlh[,2], type="bar", marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('Read Length: %{x}','<br>Count: %{y}','<extra></extra>'))
postlhp <- postlhp %>% layout(yaxis = list(title = "Count"))
postlhp <- postlhp %>% layout(xaxis = list(title = "Read Length"))
postlhp
```

<br>
<div class="rectangle"><h2 style="color:white">&nbsp;&nbsp;Clustering Statistics</h2></div>
<br>

### Number of nOTUs per clustering percentage

<br>

```{r nuc_number, echo=FALSE}
nn=read.csv(params$number_per_percentage_nucl, header=TRUE)
#nn=read.csv("number_per_percentage_nucl.csv", header=T)
nnp <- plot_ly(nn, x=nn[,1], y=nn[,2], type="scatter", mode = 'lines+markers', marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('ID%: %{x}','<br>Number of nOTU: %{y}','<extra></extra>'))
nnp <- nnp %>% layout(yaxis = list(title = "Number of nOTU"))
nnp <- nnp %>% layout(xaxis = list(title = "Clustering ID %"))
nnp
```

<br>

### Number of pOTUs per clustering percentage

<br>

```{r prot_number, echo=FALSE}
pn=read.csv(params$number_per_percentage_prot, header=TRUE)
#pn=read.csv("number_per_percentage_prot.csv", header=T)
pnp <- plot_ly(pn, x=pn[,1], y=pn[,2], type="scatter", mode = 'lines+markers', marker=list(color='#088da5', line=list(color = 'black',
               width = .1)), hovertemplate = paste('ID%: %{x}','<br>Number of pOTU: %{y}','<extra></extra>'))
pnp <- pnp %>% layout(yaxis = list(title = "Number of pOTU"))
pnp <- pnp %>% layout(xaxis = list(title = "Clustering ID %"))
pnp
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
